---
title: "Global Terrorism Adjusted by Population"
author: "Nan C. Pond"
date: "July 18, 2016"
output: html_document
---

Okay, here's our starting question:
Loosely inspired by a Washington Post article:
"https://www.washingtonpost.com/graphics/world/the-scale-of-terrorist-attacks-around-the-world/"

and a very apt comment:
"Europe + USA doesn't even account for 1/6th the world's population, while Asia alone makes up nearly 2/3rds.
I think if this was population adjusted it would move in the correct direction.
If it was Muslim population adjusted it would look even more interesting, but I guess it is not politically correct to point that out..."

Coupled with a personal burden, which is that the Western media [and dare I say, the American public?] seem to only care when white people are hurt..

What we're wondering is, is there a difference in the frequency? intensity? destructive impact of? 'terrorism' by nation, by region of the world, by relative proportion of religious practice, by GDP, by some other factor?
This is a long term question. Today's question is simply, **does a population-adjusted or identified-as-Muslim-population-adjusted comparison of the scale of casualties in terrorist attacks look [i.e. visualized, graphic form, no tests of significance] different from the raw numbers put forth by the Washington Post**

```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
