---
title: "Global Terrorism Adjusted by Population"
author: "Nan C. Pond"
date: "July 18, 2016"
output: html_document
---

Okay, here's our starting question:
Loosely inspired by a Washington Post article:
"https://www.washingtonpost.com/graphics/world/the-scale-of-terrorist-attacks-around-the-world/"

and a very apt comment:
"Europe + USA doesn't even account for 1/6th the world's population, while Asia alone makes up nearly 2/3rds.
I think if this was population adjusted it would move in the correct direction.
If it was Muslim population adjusted it would look even more interesting, but I guess it is not politically correct to point that out..."

Coupled with a personal burden, which is that the Western media [and dare I say, the American public?] seem to only care when white people are hurt..

What we're wondering is, is there a difference in the frequency? intensity? destructive impact of? 'terrorism' by nation, by region of the world, by relative proportion of religious practice, by GDP, by some other factor?
This is a long term question. Today's question is simply, **does a population-adjusted or identified-as-Muslim-population-adjusted comparison of the scale of casualties in terrorist attacks look [i.e. visualized, graphic form, no tests of significance] different from the raw numbers put forth by the Washington Post?**

## Datasets

Initial data inputs necessary are information on terrorist attacks, preferably already with information about the countries in which they occurred, and global population and dominant religion data by nation. A coarse dataset at the national level will be sufficient for time being, though of course sub-national level data would reveal more.

A quick Google search points us at:

National Consortium for the Study of Terrorism and Responses to Terrorism (START). (2016). Global Terrorism Database [Data file]. Retrieved from https://www.start.umd.edu/gtd on 18 July 2016.

Thanks, UMD.

Population data will be sourced from:

and ISO country codes:
https://datahub.io/dataset/iso-3166-1-alpha-2-country-codes/resource/9c3b30dd-f5f3-4bbe-a3cb-d7b2c21d66ce

```{r setup}
library(dplyr)
library(ggplot2)
library(readxl)

folderPath <- '/Users/Nan/dock/'

ter <- read_excel(paste0(folderPath,                      'globalterrorismdb_0616dist.xlsx')) %>%
  filter(iyear >= 1998) %>%
  select(iyear, country, region, country_txt, region_txt, nkill, nkillter)

pop_reg <- read_excel(paste0(folderPath,
                         'World Religion Dataset - National Religion Dataset.XLS'))

countryCodes <- read.csv(paste0(folderPath,
                                'iso_3166_2_countries.csv'))


```

Okay, we've got three datasets. We're going to filter everything to 1998.

```{r terFirst, echo=FALSE}
ter98 <- ter %>%
  filter(iyear >= 1998) %>%
  select(iyear, country, region, country_txt, region_txt, nkill, nkillter) %>%
  #then we want to make sure we're isolating 'innocents' deaths vs. attackers 
  mutate(nkillVictim = nkill - nkillter)

#Remove giant dataset
rm(ter)

roughTer <- ter98 %>% group_by(region_txt, iyear) %>%
  summarize(totVictims = sum(nkillVictim, na.rm = T))

ggplot(aes(x = iyear, y = totVictims, color = region_txt), data = roughTer) +
  geom_point() + theme_bw() + 
  facet_wrap(~ region_txt)

```

Pause and recall that we're looking at a brightly-colored plot of a heart-breaking number of `r sum(roughTer$totVictims)` individuals killed in terrorist attacks throughout the world.    

*Proverbs 3:31 Do not envy a man of violence And do not choose any of his ways. #

We'll get the population set to a similar place:

```{r }

majReg <- pop_reg %>%
  filter(YEAR == 2010) %>%
    select(contains('GEN'), -contains('PCT')) %>%
  mutate_all(funs(gsub(',', '', .))) %>%
  mutate_all(funs(as.numeric(.))) %>%
  colSums(., na.rm = T) %>%
  as.data.frame(.) 

majReg$Faith <- row.names(majReg)
names(majReg) <- c('subPop', 'Faith')
 
majReg <- majReg %>%
  arrange(desc(subPop)) %>%
  slice(1:5)
#This gives us the top 5 religions in the world, as of 2010

pop <- pop_reg %>%
   mutate(YEAR = gsub(',', '', YEAR)) %>%
  mutate(YEAR == as.numeric(YEAR)) %>%
   filter(YEAR == 2010) %>%
   rename(iyear = YEAR) %>%
   mutate(POP = gsub(',', '', POP)) %>%
   mutate(popNum = as.numeric(POP)) %>%
    select(iyear, one_of(majReg$Faith), ISO3, NUMISO, COUNTRY, popNum) %>%
  group_by(iyear, ISO3, NUMISO, COUNTRY) %>%
  mutate_each(funs(gsub(',', '', .))) %>%
   mutate_all(funs(as.numeric(.))) %>%
  ungroup() 

popRat <- pop %>%
  select(-iyear) %>%
  group_by(ISO3, NUMISO, COUNTRY, popNum) %>%
  mutate(chrisProp = CHRSGEN / popNum,
         islProp = ISLMGEN / popNum,
         hindProp = HINDGEN / popNum,
         syncProp = SYNCGEN / popNum,
         budProp = BUDGEN / popNum) %>%
  ungroup() %>%
  mutate(country_txt = COUNTRY) 

popMerge <- popRat %>%
  left_join(ter98)

```

First we can try to approximate that Washington Post figure:
```{r baseFigs}

popTab1 <- popMerge %>%
  group_by(iyear, region_txt) %>%
  summarize(totDeath = sum(nkillVictim, na.rm = T)) %>%
  ungroup() %>%
  mutate(totDeathAdj = ifelse(region_txt %in% c('Western Europe',
                                             'North America'),
         totDeath, -totDeath))

ggplot(aes(x = iyear, y= totDeathAdj,
           color = region_txt), data = popTab1) + 
  geom_point() + geom_point() +
  theme_bw()

ggplot(aes(x = iyear, y= totDeathAdj,
           color = region_txt), data = popTab1) + 
  geom_point() + geom_point() +
  theme_bw() + geom_line()

```

Next, we want to population adjust this. As a percentage of national population, how many victims were there?
We're going to use the 2010 population as baseline here.

```{r popAdjust}

popAdjust <- popMerge %>%
  group_by(iyear, region_txt) %>%
  summarize(totDeath = sum(nkillVictim, na.rm = T),
            totPop = sum(popNum, na.rm = T)) %>%
  ungroup() %>%
  mutate(propKilled = totDeath / totPop) %>%
  mutate(prop_of_popKilled = ifelse(region_txt %in% c('Western Europe',
                                             'North America'),
         propKilled, -propKilled))


ggplot(aes(x = iyear, y = prop_of_popKilled,
           color = region_txt), data = popAdjust) + 
  geom_point() + geom_point() +
  theme_bw()

ggplot(aes(x = iyear, y= prop_of_popKilled,
           color = region_txt), data = popAdjust) + 
  geom_point() + geom_point() +
  theme_bw() + geom_line() +
  facet_wrap(~ region_txt)

```
Hey look, North America has disappeared. Violence in Central America and Central Asia jumps out, and terrorism in Western Europe and North America disappears. Africa

Finally, we'l
```{relAdjust}



```


*Micah 6:8 He has told you, O man, what is good; and what does the LORD require of you but to do justice, and to love kindness, and to walk humbly with your God.